---
title: "WUP_04"
author: "Riley Wheeler"
date: "Summer 2022"
output:
  html_document:
    df_print: paged
---

```{r include=FALSE,echo=FALSE}
require(knitr)
require(tidyverse)
require(tigerstats)
require(mosaic)
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

The data is in the file GSScsv.csv and came from the 2018 General Social Survey.  The question to answer is to resolve this hypothesis.  
$H_0$ There is no difference between the average score on the WORDSUM test and level of schooling DEGREE  
$H_a$ There is a significant difference in average score on the WORDSUM test between at least two levels of DEGREE  

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

```{r}
gssDataframe=read.csv("GSScsv.csv")
```
WORDSUM is a Numerical variable and DEGREE is a Categorical variable.

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->
```{r}
boxplot(WORDSUM~DEGREE,data=gssDataframe)
```

From the boxplot, it appears that students less than high school have a median less than that of students with a highschool and bachelors degree.

### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
favstats(WORDSUM~DEGREE,data=gssDataframe)
```  
From this table, it appears that the median of the Graduate students is the greatest with an 11. There is a difference in the ability in the WORDSUM category.  

## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->
```{r}
model1 <- aov(WORDSUM~DEGREE,data=gssDataframe)
summary.aov(model1)
```  
Since the P value is .000236 < .05, we must reject the H_0. 

```{r}
#we specify the margins starting at the bottom and proceeding clockwise - thus 8.7 is the left side margin to hold the names of the factors
par(mar = c(3, 15, 2, 2))
plot(TukeyHSD(model1),las=2)
```

Based on the graph, there is a difference between High school graduate and non high school Graduate because these categories do not cross or touch zero.  

# Discussion and Conclusion

There is a significant difference between education degrees and WORDSUM scores. The higher levels of education, being high school graduates and beyond are much more likely to score well. This correlation between the variables means we must reject the null hypothesis that there is no difference between the average score on the WORDSUM test and level of schooling DEGREE.

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->
